Reproducible  Research : Peer Assessment 1
===========================================

##Loading and preprocessing the data
```{r setoptions}
library(knitr)
opts_chunk$set(echo=TRUE)
```

```{r loading}
library(dplyr)
library(ggplot2)
activityData=read.csv("activity.csv")
activityData$date=as.Date(activityData$date,format="%Y-%m-%d")
activityData$stepsNA=is.na(activityData$steps)
activityData_withoutNA=subset(activityData,stepsNA==FALSE)
activityData_withoutNA=activityData_withoutNA[,-4]
activityData_withNA=subset(activityData,stepsNA==TRUE)
activityData_withNA=activityData_withNA[,-4]
```




##What is mean total number of steps taken per day?

```{r mean_steps}
grouped_activityData=group_by(activityData_withoutNA,date)
summarized_activityData=summarize(grouped_activityData,steps=sum(steps))
hist(summarized_activityData$steps, breaks=20, main="Total Number of Steps Taken Each Day", xlab="Total Steps taken daily")
meanSteps=mean(summarized_activityData$steps)
medianSteps=median(summarized_activityData$steps)

```

* The mean of total number of steps taken per day is `r meanSteps`

* The median of total number of steps taken per day is `r medianSteps`





##What is the average daily activity pattern?

```{r daily_activity_pattern}
grouped_activityData=group_by(activityData,interval)
summarized_activityData=summarize(grouped_activityData,steps=mean(steps,na.rm=TRUE))
plot(summarized_activityData$interval,summarized_activityData$steps,type="l",xlab="Interval",ylab="Average steps of all days")
maxInterval=summarized_activityData[summarized_activityData$steps==max(summarized_activityData$steps),][1]
maxSteps=summarized_activityData[summarized_activityData$steps==max(summarized_activityData$steps),][2]

```

* The 5 minute interval `r maxInterval` contains the maximum number of steps ie. `r maxSteps` steps, on average across all days in the dataset.





##Imputing missing values

```{r imputing_NAs}
missingVals=table(complete.cases(activityData))[1]
grouped_activityData=group_by(activityData,interval)
summarized_activityData=summarize(grouped_activityData,steps=mean(steps,na.rm=TRUE))
m=merge(activityData,summarized_activityData,by="interval",all=TRUE)
for(i in 1:nrow(m)){
  if(is.na(m$steps.x[i])){
    m$steps.x[i]=m$steps.y[i]
  }
}
m=m[,-4]
new_activity_data=m
grouped_activityData=group_by(new_activity_data,date)
summarized_activityData=summarize(grouped_activityData,steps=sum(steps.x))
hist(summarized_activityData$steps, breaks=20, main="Total Number of Steps Taken Each Day", xlab="Total Steps taken daily")

newMeanSteps=mean(summarized_activityData$steps)
newMedianSteps=median(summarized_activityData$steps)

```

* There are `r missingVals` missing values in the dataset ie. total number of rows with NAs.

* The strategy used to impute missing values in the step variable is to replace the NAs with the mean for that 5 minute interval. 

* The mean of total number of steps taken per day is `r newMeanSteps`
The median of total number of steps taken per day is `r newMedianSteps`

* The mean after imputing NAs is the same as before imputing NAs but the median has slighly increased by imputing missing values.





##Are there differences in activity patterns between weekdays and weekends?

```{r weekend_weekday_pattern}
new_activity_data$weekdays=weekdays(new_activity_data$date)
new_activity_data$day<-ifelse((weekdays(new_activity_data$date) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")), "Weekday","Weekend")
grouped_activityData=group_by(new_activity_data,interval,day)
summarized_activityData=summarize(grouped_activityData,steps=mean(steps.x))

g=ggplot(summarized_activityData,aes(interval,steps))
g+geom_line(aes(color=day))+facet_grid(day~.)
```